% In order to understand the concept of horizontal security, we must first examine
% the current state of security and its goals. For the purposes of this work, we
% will examine security through the traditional scheme of Confidentiality,
% Integrity and Availability.

% XXX: I'd like to flip "individual' to entity, as it feels that we could be
% talking about bots/platforms/processes and more in this horizontal security
% fashion
% XXX: this also makes me wonder wether we want to bring the cyborg concept in
% a playful way. It is not fundamental to the argument, but in a way we are
% identifying these non-humans as part of the collective.
Central to the issue of Horizontal Security are the notions of
\emph{authorization} (i.e., who is allowed to access a particular resource) and
\emph{authentication} (i.e., determining who a given individual is). If
an authorized party reads, modifies, or destroys information, the definitions
imply that there is no breach. Thus, whether or not the security of an
organization has been breached depends highly on whether or not an entity is
authorized to perform an action. This decision as to whether or not an entity is
authorized is heavily informed by an organization's security policy.

% Thus, from these definitions we get that confidentiality is the protection of 
% information from unauthorized disclosure, integrity is the protection of
% information from unauthorized modification or destruction, and availability is
% the protection from the disruption of access to or use of information or an
% information system.
%

% Here, I tried to remove the old text relating to policy vs implementation and
% instead I'm trying to highlight the highly hierarchical notion of regular AC
% systems.  As a consequence, I'm leaving this
% policy/enforcement/implementation triad to arise naturally later in the text
However, as is natural with the notion of authorization, two fundamental pieces
must exist: authorization \emph{policies} that represent collective-based
authorization, and an access control system that can \emph{enforce} such
policies. Though \emph{authentication} is also a vital piece of this process,
in this work we assume that \emph{authentication} is performed by another system
and do not consider it further in this work.

\subsection{Hierarchy and Access Control}
\label{sec:definition}

Access control is commonly both encoded and enforced in a top-down fashion. The
CISO of an organization holds institutional power to define access control
policies within an organization. Further, without loss of generality,
lower-level administrators may have some decision making power in how to define
subsets of a policy, yet, they typically do not have decision making power to
change the direction of the policy completely. 

This concept typically holds true not just with access control and
authorization, but with all aspects of an organization's security policy.
Decisions about how infringements are dealt with or the scope of the policies
are also made in this manner. Decisions about how these policies are enforced,
and what security mechanisms are used, are also made up top, with entities at
the bottom left to enact the desires passed down to them.

We see, then, that hierarchy introduces itself in multiple dimensions of
security. These dimensions include the different decision-making processes of
security. However, in democratic institutions security can be multilateral
\cite{multilateral}, and thus hierarchical security decision making is not
appropriate. Instead, it is possible that horizontal security may be achieved by
shifting policy-making powers, implementation decisions, enforcement
powers, and other dimensions of security away from individuals and towards
democratic, participatory processes\footnote{Further questions arise when we
consider the structures of organizations with flexible memberships, who operate
on principles of open participation (plein-air citizens assemblies, etc.), but
these will be explored in future work}.

%\begin{definition}[Horizontality of Policy]
%Consider a set of stakeholders $ S \neq \emptyset$. There exists a set of
%security policymakers, $ p $ such that $ p \subseteq S $ and $p \neq 
%\emptyset$. We say that the horizontality of policy making $h_{policy}$ of the
%organization is defined as $ h_{policy} = \frac{|p|}{|S|} $.\\
%\noindent\rule{.48\textwidth}{0.4pt}
%\end{definition}
%
%\begin{definition}[Horizontality of Implementation]
%Consider a set of stakeholders $ S \neq \emptyset$. There exists a set of
%security implementers, $ i $ such that $ i \subseteq S $ and $i \neq 
%\emptyset$. We say that the horizontality of implementation $h_{implementation}$
%of the organization is defined as $ h_{implementation} = \frac{|i|}{|S|} $.\\
%\noindent\rule{.48\textwidth}{0.4pt}
%\end{definition}
%
%\begin{definition}[Horizontality of Enforcement]
%Consider a set of stakeholders $ S \neq \emptyset$. There exists a set of
%security enforcers, $ e $ such that $ e \subseteq S $ and $e \neq \emptyset$.
%We say that the horizontality of enforcement $h_{enforcement}$ of the
%organization is defined as $ h_{enforcement} = \frac{|e|}{|S|} $.\\
%\noindent\rule{.48\textwidth}{0.4pt}
%\end{definition}
%
%\begin{definition}[Horizontality of Security]
%We say the horizontality of an organization's security $h$ is defined as $h = 
%h_{policy} \times h_{implementation} \times h_{enforcement}$.\\
%\noindent\rule{.48\textwidth}{0.4pt}
%\end{definition}

% FIXME: I'm not entirely sure if this matters a lot for the central point
% we are not trying to measure horizontality here anymore. Instead, we want
% to show how a horizontal system can be managed.
% SOLTUION: I think we just remove it then, right? Measuring is not the goal of
% this paper.
%One important aspect of this concept is the fact horizontality requires
%different dimensions of security -- policy making, implementation, enforcement,
%and more. If one area is highly hierarchical and centralized, it has a profound
%effect on the hierarchy and centralization of the organization as a whole. For
%example: if the implementation of a security policy (through choosing security
%mechanisms, implementing firewall rules, etc.) is heavily centralized and
%hierarchical, then the member(s) of the organization who implement the policies
%can decide how to implement security mechanisms to ensure policies are
%followed, or worse still, whether to implement any security mechanisms at all.
%This centralization leads to an imbalance of power similar to the one discussed
%by Kavada regarding activism and social media~\cite{kavada2020counterpublics}.

% Fixme: this brings up the question of *identity*. Entities are multi-phrenic,
% how can we reconcile access to resources when everybody can wear many hats
% within the organization. Signed: the chair of shitposting society at 
% HorizontalSecurityPaper enterprises. I'm actually a little bit more interested
% in hearing about this than the binary-ness of this (and I think they are
% deeply connected)
% FIXME: I struggled with this part. Please reread it and see if anything is
% worth salvaging. After reading it many times I think it's terrible. If you
% think you can save the idea, awesome, but my recomendation would be to just
% take this paragraph out.
% NEW (KG):
%When shifting to these participatory processes, identity still remains a central
%question in access control decisions. When determining whether or not to grant
%access, the collective will consider which entity or entites they wish to grant
%access to, taking into account many attributes included competency, trust, and
%current power dyanmics within the collective. Thus, though other access control
%schemes like capabilities can create identity-agnostic protocols, collective
%based access control systems require this information to be present.
% DELETED:
%Another implication of the fact that the horizontality of security is
%multi-dimensional is that horizontality is not a binary trait, but rather a
%spectrum between entirely dictatorial and fully horizontal. In practice, most
%organizations likely do not sit at either extreme of this spectrum, or even
%occupy a fixed space on this spectrum. Rather, organizations are likely to have
%dynamic levels of horizontality. In traditional, hierarchical organizations,
%security is likely more skewed towards a lower amount of horizontality, and is
%likely to not be too dynamic. Decisions are made by managers, who may or may not
%consider the input of their employees when making decisions. In worker
%cooperatives,  activist groups, and other traditionally less hierarchical
%organizations, security is more likely skewed towards a higher amount of
%horizontality, and may be much more dynamic in nature. An interesting first work
%towards understanding horizontal security may exist in classifying, comparing,
%and contrasting the current horizontality of security in different types of
%organizations. This idea, and other potential future works, are further examined
%in Section~\ref{sec:futurework}.

% TODO: Include here the lack of ability to be horizontal when hierarchical
% tools are used, and vice-versa, the lack of ability to be hierarchical when
% horizontal tools are used. Can we develop technology that allows us to do
% both when they are needed?
However, organizations that attempt to create their access control policies
horizontally may run into problems when they move to implement them. Currently,
access control systems force the creation of a hierarchy. In access control
systems such as MAC, DAC, or RBAC, certain entities in the organization have the
power to implement the access control rules that were democratically created by
the organization. Thus, if the individuals who have the ability to enforce the
rules decide not to, the newly created access control policies become
ineffective. This forms a hierarchy with the entities capable of enforcing
access control sitting at the top. Similarly, if an organization were to attempt
to temporarily act hierarchically and to implement that using tools
that assumed horizontality, this may cause the hierarchy to lose any potential
efficiency benefits. Thus, a question arises: can we develop technology that
allows an organization to be flexible and dynamic in its horizontality, being
participatory or hierarchical depending on the needs at hand, without
jeopardizing the ability to return to horizontality when desired?

\subsection{Previous Attempts at Horizontal Security}
\label{sec:previous}
Having explored the principles that unlock horizontality, we can move on to
creating specific systems that allow entities to organize in such a way.
Achieving horizontality is not trivial. In fact, many projects and
organizations have attempted to do so with varying degrees of success. In this
section, we argue that these failings are a consequence of a missing building
block: Collective-based Access Control (or COLBAC).

% XXX: I removed this paragraph below because it was repetitive imvho.
% In Section~\ref{sec:definition} we discussed horizontality of security as being
% multidimensional and dynamic, and relating to the different decision making
% processes that an organization has around security. Thus, we see that
% horizontality is related to security policy decisions, as the participation in
% these decision making processes informs what power individuals in the
% organization have to say how they wish to secure their operations.

% FIXME: ok, I get it, we've been hammering on the verticality of organizations
% and their tools, do we need more?
% However, this does not mean that the influence of hierarchy does not affect the
% tools and techniques used in Cybersecurity, and that some tools and techniques
% do not imply or impose hierarchy. For example, most operating systems
% tend to be built around the idea of the singular administrator or a small set of
% administrators setting up the system. This idea makes sense from a  hierarchical
% organization's perspective. After all, if the individuals higher in the
% hierarchy have more privileges, that simply reflects their decision-making 
% power within the organization. From a horizontal security perspective, however,
% this method of system administration is detrimental. If a singular administrator
% decides that they do not like the policy decisions of the organization, they may
% be able to use their administrative power to stop the organization from taking
% the democratically selected action. When this occurs, there may not be another
% individual with more privileges to override this decision. Thus, the very
% concept of an administrator with the power to shut out other administrators is a
% threat to a horizontally controlled organization.

When laying out known horizontal software systems, perhaps the most
widely-known example is Bitcoin~\cite{nakamoto2019bitcoin}. This cryptocurrency
was developed to allow for financial transactions between two individuals
without the need for a trusted third party to detect and prevent double
spending. By design, the network prevents the modification of past blocks of
the Blockchain unless there exists collusion of a large group of entities with
computing~\cite{bonneau2018hostile} (or to a lesser extent,
network~\cite{heilman2015eclipse}) resources. This decision was made
specifically to make it difficult for malicious actors in the network to
perform malicious activities, such as stealing bitcoin or deleting past
transactions. The implication of this design is that Bitcoin cannot be
centrally regulated~\cite{tu2015rethinking}, at least in theory. Thus, a core
element of cryptocurrencies, the consensus protocol is an instantiation of a
democratically-based access control policy. %(this is indeed more obvious for
%proof-of-stake Blockchain systems).

Yet there exist many shortcomings in the Bitcoin/cryptocurrency model. First
is identifying \emph{who} is part of the collective: are only miners
stakeholders, or are users stakeholders as well, are developers stakeholders?
This question raises the fundamental limitation of Bitcoin. Although it can be
considered ``network-horizontal'' it does not provide  horizontality mechanisms
for other elements of the ecosystem. This is apparent when considering
historical events in Blockchain-based systems, such as the DAO hack~\cite{EthDAOHack},
the Bitcoin Classic fork~\cite{bitcoinForks} and more. Indeed, it seems that developers of
a Blockchain, those blessed with public visibility (or even just commit access)
can have more impact on the nature of the Blockchain than a large mining
cartel.
% All miners participate in the decision to perform a
% modification on the Blockchain, whether they vote yes or no. If we consider
% users to also be stakeholders, however, we see that this system is no longer as
% horizontal. Users do not get to participate in the decision of whether or not a
% transaction gets removed or modified. 

Another issue with this approach is its competitive nature. Rather than
cooperating miners attempting to create, implement, and enforce transaction
policies, we have competing miners interested in their own profit. One
implication of this competitive focus is that there is no mechanism to vote in
Bitcoin; rather, what is referred to as ``regulation'' in this work is seen by
the Bitcoin community as an attack. However, the design property of requiring
a majority of participating members to agree in order to make a change on the
system may still be useful to inspire future horizontal system designs. In the
case of cryptocurrencies, the competitive design tends to centralize the network
over time. Bitcoin's Proof of Work algorithm tends to favor those who can afford
highly specialized hardware, eventually leading to a centralization of the
network into the hands of those who are wealthy enough to control large amounts
of machines specifically made for mining Bitcoins. Proof of Stake, another
consensus algorithm for Blockchain, bypasses this middle step and directly
benefits those who control larger amounts of monetary stake in the network.

Other, more voluntary (and less profit-driven) designs also attempt to
achieve horizontality. Perhaps the most well known example of a security system
meant to be horizontal by design is the PGP Web of Trust (WoT) design. This
ecosystem attempts to allow entities to authenticate (i.e., tie public key
material to emails, names, or identities) and exchange signed or encrypted
messages without a central authority. This is done through a transitive trust
process: known-good entities can sign trust relationships with other entities
and various WoT algorithms will compute resulting trust levels based on a
personal security policy that ranges from marginal trust to full trust. There is
however no global notion of policing or enforcement in the PGP ecosystem. As
such, it is impossible for entities in a group to agree access control policies
in a distributed ecosystem.

% This is a rather tangential problem, that lies within GPG design, not PGP
% However, this scheme suffered from a large
% number of usability and user experience problems~\cite{whitten1999johnny}.
% However, if a method like web of trust or derived from web of trust could be
% improved to be more usable, it could serve as a building block for applications
% and systems that take a horizontal approach to security.

% XXX: This last paragraph seems to diverge from the message a bit. It would
% be worth it to reframe it to be more along the lines of "other systems that
% attempt to solve the problem or part of the problem but still don't arrive
% at a complete solution."
In addition to the Blockchain approach used by Bitcoin, and to the web of trust
used by PGP, many other attempted approaches towards horizontal security already
exist. These include other consensus algorithms such as Practical Byzantine
Fault Tolerance (PBFT)~\cite{castro1999practical}, Paxos~\cite{lamport2019part}, 
Raft~\cite{ongaro2013search}, and more. However, most of these protocols do not
allow for flexible horizontality; rather, they have a set percentage of
entities in the system that must agree. More, these systems are more concerned
with fault tolerance rather than access control. However, using these solutions
for the distribution of horizontal security technologies may be beneficial in
future projects.

In addition to consensus algorithms, secret sharing schemes like Shamir Secret
Sharing~\cite{shamir1979share},
Blakley's Secret Sharing~\cite{blakley1993linear}, and Verifiable Secret 
Sharing~\cite{chor1985verifiable} schemes can be used for distributed secret
management in organizations. This allows for individuals to recover access to a
resource if a set fraction of the participants agree and submit their secret
share. However, these approaches are inefficient and are also rigid in their
horizontality.

Other solutions to small horizontal security problems exist. Applications like
OAuth~\cite{leiba2012oauth} can be used to grant minimal and ephemeral access to
organization accounts to members of the organization who need it, without
revealing a password. However, someone in the organization still must control
the password, causing hierarchy issues. Capability~\cite{dennis1966programming}
based schemes and systems can be used to allow for transfer of control over
resources in a more flexible way, but is not sufficient alone: a collective
authorization scheme is required to determine whether or not to issue capability
tokens.

Previous attempts of expanding access control to fill some of these gaps have
been made, but each of them has its flaw. Most work centers around the idea of
multi-ownership access control or multi-party access control (MPAC), where
different files in a system are owned by multiple different users who need to
agree on who may see a given resource and who may not. These schemes are usually
specifically designed for online social networks (OSNs), and generally use
either majority voting
(\cite{Hu_Ahn_Jorgensen_2013,Ilia_Polakis_Athanasopoulos_Maggi_Ioannidis_2015},
etc.) secret sharing techniques 
(\cite{Ilia_Carminati_Ferrari_Fragopoulou_Ioannidis_2017}, etc.), or conflict
resolution protocols
(\cite{Carminati_Ferrari_2011, Bruns_Huth_2011, Rao_Lin_Bertino_Li_Lobo_2011}, 
etc.) to determine with whom files should be shared. Other approaches rely on
the relationships between the owners of a resource and other users within a
community or (OSN) to determine who gets to access a file. However, these
approaches are different from ours in many ways. Primarily, none of the MPAC
solutions focus on flexible horizontality: instead, they assume either majority
voting or some form of conflict resolution, rather than begin
flexible enough to cover multiple forms of horizontal structure. More, these
systems focus on specific issues related to OSNs, such as privacy rights and
concerns for photo sharing, and as such these systems emphasize read decisions
over write or execute decisions, making them insufficient for the purpose of
access control and authorization on a general-purpose system. Finally, previous
attempts to fill this gap consider resource sharing and authorization within a
complex community where individuals who belong to the community own files and
make decisions regarding those files. They do not consider the problem of access
control and authorization in the name of the community or collective as a whole,
as our solution does.

% XXX: We need an awesome one sentence to transition over to the idea of COLBAC
% here.
To address these gaps we present COLBAC, a collective based access control
system to provide the fundamental basis for more democratic and participatory
security software and systems.
%\input{tex/Colbac}

% Access Control:
% La-Padula, Biba, strictly hierarchical
% DAC based on individualistic regulation of files, overwritable by admin on
%    most systems.
% MAC or RBAC seem to be desirable for horizontal security, but how do we update
%    policies?

% Bitcoin:
% attempt at horizontal, democratic control of currency.
% Regulatable with 51%.
% suffers from centralization as it goes on due to its competative nature.

%\begin{enumerate}
%\item Case study: Code review.
%\item Case study: Peer review.
%\end{enumerate}

%Conclusion: Security is purely horizontal when the policies and enforcement are
%handled via direct democratic vote. Horizontality is a spectrum, ranging from
%hierarchical (one person with decision making power) to purely horizontal.
%Examples of spots on the spectrum include:
%\begin{itemize}
%\item direct democracy / consensus cooperatives
%\item Random review (i.e. code review)
%\item random/elected representatives w/ early revocation
%\item random/elected representatives w/o early revocation
%\item appointed representatives
%\item hierarchical
%\end{itemize}

%Very few spots on this spectrum are represented in current techniques and
%techologies. Mostly the hierarchical approach is developed because of its
%continued use in business and government. As that changes, cybesecurity needs to
%be ahead of the curve and create technologies for the organizations and
%businesses that sit on the other portions of this spectrum.
