Collective management of resources opens an organization to a collection of both
internal and external threats. As such, it is expected that attackers may be
outsiders impersonating members of a community as well as disgruntled insiders
that are trying to subvert the system. We expect, however, that at least a
simple majority of users will act honestly, and will attempt to use the system
to  represent the collective goal of the broader community.

However, we also expect attackers to attempt to inject accounts in attempts to
dilute the democratic process (e.g., a Sybil attack~\cite{sybil1, sybil2}).
While these types of attacks are often performed in the wild~\cite{197120},
we assume this can be solved by a strong authentication scheme, implemented as
each organization sees fit, and thus we consider this outside of our threat
model. One example of how this can be solved separately from the authorization
is the membership authentication of the Debian project, for which membership is
often gated through physically-present PGP key signing processes, forming a
strong authentication method. This has been widely documented in work by
Coleman~\cite{coleman-oss}. 

Malicious users may abuse other elements beyond raw voting power. They could,
for example, gather a group colluding users, who could abuse time-zones (e.g.,
making decisions and calling votes when many other people are not online),
system availability (e.g., network outages, net-splits) to carry out exceptional
votes. This can likely be mitigated by ensuring the vote parameters are adjusted
to the nature of the community (e.g., by adjusting simple majority to relative
majority or set a necessary threshold for quorum).
